### 2023-06-20 test_minted.tex

\documentclass{article}              % for review, blinded

\usepackage[most]{tcolorbox}
\usepackage{minted}
\tcbuselibrary{minted} % tcolorbox minted library, required to use the "minted" tcb listing engine (this library is not loaded by the option [most])
%\usepackage{minted} % Allows input of raw code, such as Python code

\begin{document}
	\begin{tcolorbox}
	\begin{minted}[gobble=1]{python}
	print("hello, world")
	\end{minted}
\end{tcolorbox}

\begin{tcolorbox}[left=1mm]
	this is code
\end{tcolorbox}
\end{document}